---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---

---
title: "Peer Graded Assignment - Course Project 1"
output: html_document
---

## Loading and preprocessing the data

As a part of preprocessing I converted date from character to Date Format
```{r , echo=TRUE}
setwd("F:\\Data Science\\My Codes and Assignments\\Coursera\\Reproducible research\\week 2\\Data")
DT <- read.csv("activity.csv")
DT$date <- as.Date(as.character(DT$date))
```

## What is mean total number of steps taken per day?

Followed the following steps:
1) I omitted the null values for this part
2) Aggregated the number of steps for each day
3)plotted the histogram 

```{r, echo=TRUE}
library(dplyr)
DTcomplete <- na.omit(DT)
summarisedDT <- DTcomplete %>% group_by(date) %>% summarise(Tot.Steps = sum(steps))
plot(summarisedDT$date,summarisedDT$Tot.Steps,type = "h",xlab = "Date",ylab = "No Of Steps",main = "No of Steps Each Day")
```


Mean of No of steps per day:
```{r, echo=TRUE}
mean(summarisedDT$Tot.Steps,na.rm = TRUE)
```

Median of No of steps per day: 
```{r, echo=TRUE}
median(summarisedDT$Tot.Steps)
```

##What is the average daily activity pattern?




```{r, echo=TRUE}
library(dplyr)
DTcomplete <- na.omit(DT)
summarisedDataComplete <- DTcomplete %>% group_by(interval) %>% summarise(Tot.Steps = sum(steps))

plot(summarisedDataComplete$interval,summarisedDataComplete$Tot.Steps,type = "l",xlab = "Interval" ,ylab = "No of Steps",main = "No of Steps per interval")

```

The Interval With the maximum number of Steps

```{r, echo=TRUE}
lastRow<- tail(summarisedDataComplete[order(summarisedDataComplete$Tot.Steps),],1)
lastRow$interval
```

##Inputing missing values

Number of missing values

```{r, echo=TRUE}
sum(is.na(DT))
```

Strategy For Finding out missing Values:

The Function Calculate Missing values calculates the missing value for that row which is he mean of the number of steps in that interval. 

```{r, echo=TRUE}

calculateMissingValue <- function(dataSet,rowIndex)
{
  row <- dataSet[rowIndex,]
  subsetOfThisInterval <- subset(dataSet,dataSet$interval == row$interval,select = c("steps"))
  return(round(mean(subsetOfThisInterval$steps,na.rm = TRUE)))
}

```


The Function ReturnMissingFilledDataSet fills creates a new dataset where missing values are filled based on the above strategy. 

```{r, echo=TRUE}
ReturnMissingFilledDataSet <- function(dataSet)
{
dataFilledMissing <- data.frame(steps = numeric(),date = numeric(),interval=numeric())
for (i in 1:nrow(dataSet)) {
  row <- dataSet[i,]
  if(is.na(row$steps))
  {
    dataFilledMissing[i,] = c(calculateMissingValue(dataSet,i),row$date,row$interval)
    next()
  }
  dataFilledMissing[i,] = c(row$steps,row$date,row$interval)
}
return(dataFilledMissing)
}


dataNAFilled = ReturnMissingFilledDataSet(DT)
dataNAFilled$date = as.Date(dataNAFilled$date, origin="1970-01-01")
```


Histogram of number of steps taken each day

```{r, echo=TRUE}
summarisedDT <- dataNAFilled %>% group_by(date) %>% summarise(Tot.Steps = sum(steps))
summarisedDT$date <- as.Date(as.character(summarisedDT$date))
plot(summarisedDT$date,summarisedDT$Tot.Steps,type = "h",xlab = "Date",ylab = "No
Of Steps",main = "No of Steps Each Day")
```

Comarison of Medians
They do differ slightly because of a wisely chosen strategy

```{r, echo=TRUE}
summarisedDTNaFilled <- dataNAFilled %>% group_by(date) %>% summarise(Tot.Steps =  sum(steps))  

DTcomplete <- na.omit(DT)
summarisedDTComplete <- DTcomplete %>% group_by(date) %>% summarise(Tot.Steps = 
sum(steps))                                                                     


median(summarisedDTNaFilled$Tot.Steps)
median(summarisedDTComplete$Tot.Steps)
```

Comparison of Means

They do not differ slightly because of a wisely chosen strategy

```{r, echo=TRUE}
mean(summarisedDTNaFilled$Tot.Steps)
mean(summarisedDTComplete$Tot.Steps)
```

##Are there differences in activity patterns between weekdays and weekends?

Creating a new factor indicating whether weekend or weekday

```{r, echo=TRUE}
library(dplyr)
weekDays <- c("Monday","Tuesday","Wednesday","Thursday","Friday")
weekEndDays<- c("Saturday","Sunday")
DT <- mutate(DT,DayOfWeek = weekdays(DT$date))
DT<- mutate(DT,DayType = ifelse((DayOfWeek %in% weekDays), "WeekDay", ifelse((DayOfWeek %in% weekEndDays), "WeekEnd", 0)))
DT$DayType <- as.factor(DT$DayType)
head(DT)
```


Panel Plot of Weekdays Vs Weekends

```{r, echo=TRUE}
dataWeekEnddays = subset(DT,DT$DayType=="WeekEnd")
dataWeekdays = subset(DT,DT$DayType=="WeekDay")
par(mfrow=c(2,1)) 
plot(dataWeekdays$interval,dataWeekdays$steps,type = "l",main = "Weekdays",ylab = "Steps",xlab = "Interval")
plot(dataWeekEnddays$interval,dataWeekEnddays$steps,type = "l",main = "WeekEnds",ylab = "Steps",xlab = "Interval")
```



